---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
```

<!-- badges: start -->
[![Travis build status](https://travis-ci.org/bcjaeger/ipa.svg?branch=master)](https://travis-ci.org/bcjaeger/ipa)
[![Lifecycle: maturing](https://img.shields.io/badge/lifecycle-maturing-blue.svg)](https://www.tidyverse.org/lifecycle/#maturing)
<!-- badges: end -->

# Imputation for predictive analytics (`ipa`)

The goal of `ipa` is to make imputation in predictive modeling workflows more straightforward and efficient. The main functions in `ipa` are

1. `brew` Create a container to hold your imputations

2. `spice` (optional) set parameters that govern the number of imputations for the given brew

3. `mash` fit models that will provide imputations

4. `ferment` impute missing values in training and (optionally) testing data 

5. `bottle` output the imputed datasets in a tibble.



## Installation

<!-- You can install the released version of ipa from [CRAN](https://CRAN.R-project.org) with: -->

<!-- ``` r -->
<!-- install.packages("ipa") -->
<!-- ``` -->

You can install the development version of `ipa` from [GitHub](https://github.com/) with:

``` r
# install.packages("devtools")
devtools::install_github("bcjaeger/ipa")
```

# Example: diabetes data

These data are courtesy of Dr John Schorling, Department of Medicine, University of Virginia School of Medicine. The data originally described 1046 participants who were interviewed in a study to understand the prevalence of obesity, diabetes, and other cardiovascular risk factors in central Virginia for African Americans. A total of 403 participants were screened for diabetes, which was defined as glycosolated hemoglobin > 7.0. Among those 403 participants, 366 provided complete data for the variables in these data.

First, we'll load some packages and set a seed for reproducibility

```{r, message = F}

library(tidymodels)
library(tidyverse)
library(ipa)

data("diabetes", package = 'ipa')

diab_complete <- diabetes$complete
diab_missing <- diabetes$missing

glimpse(diab_missing)

```

## K-nearest-neighbor imputation

K-nearest-neighbors (KNN) is a flexible and useful method for imputation of missing data. Briefly, each missing value is imputed by aggregating or randomly sampling a value from the K observations with greatest similarity to the current obervation. Conventional methods for KNN provide imputations using a single value of K, which makes it hard to identify an optimal value of K. Ideally, one would generate imputed datasets using a number of different values of K, and then use whichever K provided the most accurate imputed values or the most accurate prediction model (usually, these K are the same or similar). This is one of the things `ipa` does. 

Generally, the `ipa` workflow follows the same patterned steps. Here is an example of this workflow using k-nearest-neighbors. A more detailed explanation of the steps below is written in the `Getting started` vignette.

```{r}

nbrs_brew <- brew_nbrs(diab_missing, outcome = diabetes) %>%
  verbose_on(level = 1) %>% # makes the brew noisy
  spice(with = spicer_nbrs(k_neighbors = 1:35)) %>% # tuning parameters
  mash() %>% # creates imputed values 
  ferment() %>% # turns values into a data frame
  bottle(type = 'tibble') # turns data into tibbles, adds outcome column

nbrs_brew


```

